---
title: "psm"
output: html_document
date: "2024-06-30"
---
```{r}
install.packages("MatchIt")
install.packages("optmatch")
install.packages("Matching")

library("MatchIt")
library("optmatch")
library("Matching")
```
```{r}
set.seed(1)
a <- read.csv(file="imputed4 - imputed4.csv")

sites <- unique(a$S)
nsites <- length(sites)

```{r}
ps <- c(ma.out1[["distance"]], ma.out2[["distance"]], ma.out3[["distance"]], ma.out4[["distance"]], ma.out5[["distance"]], ma.out6[["distance"]], ma.out7[["distance"]], ma.out8[["distance"]], ma.out9[["distance"]], ma.out10[["distance"]], ma.out11[["distance"]])
data_1 <- cbind(ps,a)
data_1
write.csv(data_1, file = "nearest.csv", row.names = FALSE)
```

```{r}
ps_1 <- c(Ma.out1[["distance"]], Ma.out2[["distance"]], Ma.out3[["distance"]], Ma.out4[["distance"]], Ma.out5[["distance"]], Ma.out6[["distance"]], Ma.out7[["distance"]], Ma.out8[["distance"]], Ma.out9[["distance"]], Ma.out10[["distance"]], Ma.out11[["distance"]])
data_2 <- cbind(ps_1,a)
data_2
```



```{r}
ps_2 <- c(ma.out_1[["distance"]], ma.out_2[["distance"]], ma.out_3[["distance"]], ma.out_4[["distance"]], ma.out_5[["distance"]], ma.out_6[["distance"]], ma.out_7[["distance"]], ma.out_8[["distance"]], ma.out_9[["distance"]], ma.out_10[["distance"]], ma.out_11[["distance"]])
data_3 <- cbind(ps_2,a)
data_3
write.csv(data_3, file = "data_3.csv", row.names = FALSE)
```

optimal
```{r}
em.pair <- read.csv(file="data_3.csv") 
presf <- friedman.test(em.pair$Presentations, em.pair$REU, em.pair$EM.pair)$p.value
presf
pubf <- friedman.test(em.pair$Publications, em.pair$REU, em.pair$EM.pair)$p.value
pubf
awardsf <- friedman.test(em.pair$Awards, em.pair$REU, em.pair$EM.pair)$p.value
awardsf
```

nearest
```{r}
Em.pair <- read.csv(file="nearest.csv") 
presf <- friedman.test(Em.pair$Presentations, Em.pair$REU, Em.pair$EM.pair)$p.value
presf
pubf <- friedman.test(Em.pair$Publications, Em.pair$REU, Em.pair$EM.pair)$p.value
pubf
awardsf <- friedman.test(Em.pair$Awards, Em.pair$REU, Em.pair$EM.pair)$p.value
awardsf
```

```{r}
g= read.csv("a.pair.csv")
presf <- friedman.test(g$Presentations, g$REU, g$a.pair)$p.value
presf
pubf <- friedman.test(g$Publications, g$REU, g$a.pair)$p.value
pubf
awardsf <- friedman.test(g$Awards, g$REU, g$a.pair)$p.value
awardsf
```

optimal
```{r}
EM.diff <- em.pair$em.diff
a.diff <- g$a.diff
par(mfrow=c(1,2))
# Colors with transparency (alpha), grid lines, and refined labels
hist(EM.diff, 
     breaks=10, 
  col = rgb(0.3, 0.7, 0.4, 0.8),
     border="white", 
     main="New Pairs", 
     xlab="PS Difference", 
     ylab="Frequency",
     cex.main=1, 
     cex.lab=1, 
     cex.axis=1)
grid(nx=NULL, ny=NULL, col="lightgray", lty="dotted")

hist(a.diff, 
     breaks=10, 
    col=rgb(0.9, 0.7, 0.1, 0.7),  # Soft yellow with transparency
     border="white", 
     main="Initial Pairs", 
     xlab="PS Difference", 
     ylab="Frequency", 
     cex.main=1, 
     cex.lab=1, 
     cex.axis=1)
grid(nx=NULL, ny=NULL, col="lightgray", lty="dotted")

mtext("Figure 1: Optimal Matching", 
      side=1, outer=TRUE, line=-1, cex=0.9, font=2)  


par(mfrow=c(1,2))
hist(EM.diff, breaks=10 , col=rgb(0.2,0.1,0.7), main="New Pairs", xlab="PS difference")
hist(a.diff, breaks=10 , col=rgb(0.8, 0.4, 0), main="Initial Pairs", xlab="PS difference")

#par(mfrow=c(1,1))
#boxplot(EM.diff, a.diff,at = c(1,2), main = "Propensity Score Difference Within Each Match", names = c("new match", "initial match"),col = c( rgb(0.2,0.1,0.7), rgb(0.8, 0.4, 0) ),  notch = FALSE, border = "black",horizontal = TRUE, showmeans=TRUE )
#legend("topleft", inset=.007, title="Mean",
 #  c("new=0.6105","initial=0.8138"), fill = c( rgb(0.2,0.1,0.7), rgb(0.8, 0.4, 0)), horiz=TRUE, cex=0.8)
```

nearest neighbor
```{r}
em.diff <- Em.pair$em.diff
a.diff <- g$a.diff
# Setting up the layout for side-by-side histograms
par(mfrow=c(1,2))

# Colors with transparency (alpha), grid lines, and refined labels
hist(em.diff, 
     breaks=10, 
    col=rgb(0.1, 0.6, 0.8, 0.7),
     border="white", 
     main="New Pairs", 
     xlab="PS Difference", 
     ylab="Frequency",
     cex.main=1, 
     cex.lab=1, 
     cex.axis=1)
grid(nx=NULL, ny=NULL, col="lightgray", lty="dotted")

hist(a.diff, 
     breaks=10, 
    col=rgb(0.9, 0.7, 0.1, 0.7),  # Soft yellow with transparency
     border="white", 
     main="Initial Pairs", 
     xlab="PS Difference", 
     ylab="Frequency", 
     cex.main=1, 
     cex.lab=1, 
     cex.axis=1)
grid(nx=NULL, ny=NULL, col="lightgray", lty="dotted")

mtext("Figure 2: Nearest Neighbor Matching ", 
      side=1, outer=TRUE, line=-1, cex=0.9, font=2)  

# Smaller font for the description


par(mfrow=c(1,2))
hist(em.diff, breaks=10 , col=rgb(0.2,0.1,0.7), main="New Pairs", xlab="PS difference")
hist(a.diff, breaks=10 , col=rgb(0.8, 0.4, 0), main="Initial Pairs", xlab="PS difference")



#par(mfrow=c(1,1))
#boxplot(em.diff, a.diff,at = c(1,2), main = "Propensity Score Difference Within Each Match", names = c("new match", "initial match"),col = c( rgb(0.2,0.1,0.7), rgb(0.8, 0.4, 0) ),  notch = FALSE, border = "black",horizontal = TRUE, showmeans=TRUE )
#legend("topleft", inset=.007, title="Mean",
   #c("new=0.6105","initial=0.8138"), fill = c( rgb(0.2,0.1,0.7), rgb(0.8, 0.4, 0)), horiz=TRUE, cex=0.8)
# Setting up the layout for side-by-side histograms
```
bar plots 

```{r}
install.packages("ggplot2")
```

```{r}
library(ggplot2)
# Gender distribution
ggplot(em.pair, aes(x = Gender)) +
  geom_bar(aes(fill = Gender), width = 0.6) +  # Different colors for each gender
  scale_fill_manual(values = c("male" = "#4B8BBE", "female" = "#E6B73A")) +  # Custom colors
  labs(title = "Gender Distribution of Participants",
       x = "Gender",
       y = "Count") +
  theme_minimal(base_size = 15) +  # Increase base font size
  geom_text(stat = "count", aes(label = after_stat(count)), 
            vjust = -0.2, size = 4, position = position_dodge(width = 0.7)) +  
  theme(legend.position = "none") 
```
```{r}
ggplot(em.pair, aes(x = Race.Ethnicity)) +
  geom_bar(aes(fill = Race.Ethnicity), width = 0.7) +  
  scale_fill_brewer(palette = "Set3") +  
  labs(title = "Race/Ethnicity Distribution of Participants",
       x = "Race/Ethnicity",
       y = "Count") +
  theme_minimal(base_size = 15) +  
  geom_text(stat = "count", aes(label = after_stat(count)), 
            vjust = -0.3, size = 4, position = position_dodge(width = 0.7)) +  
  theme(legend.position = "none")  
```
```{r}
ggplot(em.pair, aes(x = Birth.Year)) +
  geom_histogram(binwidth = 1, fill = "#4B8BBE", color = "white", alpha = 0.8) +  # Use a richer purple and add transparency
  labs(title = "Distribution of Birth Years",
       x = "Birth Year",
       y = "Count") +
  theme_minimal(base_size = 15) +  # Increase base font size
  geom_text(stat = "count", aes(label = after_stat(count)), 
            vjust = -0.3, size = 4, position = position_identity()) +  # Add count labels on top of bars
  scale_x_continuous(breaks = seq(min(em.pair$Birth.Year), max(em.pair$Birth.Year), by = 1)) +  # Custom x-axis breaks
  theme(panel.grid.major = element_line(color = "gray90"),  # Lighter grid lines for better visibility
        panel.grid.minor = element_blank(),  # Remove minor grid lines
        plot.title = element_text(hjust = 0.5))  # Center the title
```

```{r}
library(ggplot2)
# Histogram of GPA
ggplot(em.pair, aes(x = GPA)) +
  geom_histogram(binwidth = 0.2, fill = "#4B8BBE", color = "white") +
  labs(title = "Distribution of GPA", x = "GPA", y = "Count") +
  theme_minimal()
```


```{r}
# Stacked bar plot for Gender and Highest Degree Level
ggplot(em.pair, aes(x = Highest.degree.level.re.written, fill = Gender)) +
  geom_bar(position = "fill") +
  labs(title = "Proportional Distribution of Gender by Highest Degree Level",
       x = "Degree Level",
       y = "Proportion") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("male" = "#4B8BBE", "female" = "#E6B73A")) +  # Custom colors for each gender
  theme_minimal(base_size = 15) +  # Increase base font size
  theme(legend.title = element_blank(),  # Remove legend title
        plot.title = element_text(hjust = 0.5))  # Center the title
```
```{r}
# Facet grid for GPA by Gender and Degree Level
ggplot(em.pair, aes(x = GPA, fill = grad.vs.undergrad)) +
  geom_histogram(binwidth = 0.2, color = "white") +
  labs(title = "GPA Distribution by Graduate vs Undergraduate", x = "GPA", y = "Count") +
  facet_wrap(~ Gender) +
  theme_minimal()
```


